IPv6 over Low-Power Wireless Personal Area Networks\\ MSc. Thesis Proposal

### 2010-01-17 trunk/doc/msc-proposal/msc-proposal.tex
proposal: corrected ieee 802.15.4

\chapter{IEEE 802.15.4}\label{ch:ieee802.15.4}
The IEEE 802.15.4 standard \cite{ieee802.15.4} specifies the physical layer and media access control for low-rate wireless personal area networks. The standard intends to offer the fundamental lower network layers of a type of wireless personal area network which focuses on low-cost, low-speed communication between tiny devices (motes). Such networks are typically limited to a personal operating space (POS) of up to 10 meters with a transfer rate of 250 kbit/s and involve little or no infrastructure.

The IEEE 802.15.4 standard distinguishes between two types of nodes, reduced-function devices (RFDs) and full-function devices (FFDs).                                   FFDs typically have more resources and may be mains powered.  Therefor, FFDs aid RFDs by providing functions such as network coordination and packet forwarding.

An 802.15.4 network may operate in either the star or the peer-to-peer topology. In the star topology devices communicate with a single central                           a personal area network (PAN) coordinator. In the peer-to-peer topology a device can communicate with any other device as long as they are in range of one another. Based on the peer-to-peer topology more complex network formations may be constructed, such as mesh networking topology. However, the standard does not define a network layer, therefor routing is not directly supported, but such an additional layer can add support for multihop communications. All devices operating on a network of either topology have unique 64 bit extended addresses. This address can be used for direct communication within the PAN, or it can be exchanged for a short 16 bit address allocated by the PAN coordinator.

The basic unit of data transport is a frame. The standard defines four frame structures: a beacon frame, used by a coordinator to transmit beacons;
a data frame, used for all transfers of data; an acknowledgment frame, used for confirming successful frame reception; a MAC command frame, used for handling all MAC control transfers. Additionally, a superframe structure may be defined by the coordinator. In such case two beacons act as superframe limits and provide synchronization to other devices. A superframe consists of sixteen equal-length slots, which can be further divided into an active part and an inactive part, during which the coordinator may enter power saving mode. Any device wishing to communicate during the contention access period (CAP) between two beacons shall compete with other devices using a slotted CSMA-CA mechanism. For applications requiring specific data bandwidth, the PAN coordinator may dedicate portions of the active superframe, which form the contention-free period (CFP).

An important aspect of 802.15.4 is its limitation on the frame size, which is specified by the frame length 7 bit field (0-127 bytes). Taking into account the frame header, which is up to 25 octets, this leaves 102 bytes for the payload of the higher layers.

The standard also defines several security services such as maintaining an access control list (ACL) and
using symmetric-key cryptography to protect transmitted frames. However, key management is not specified and must be provided by the higher layers.

%%
%	Transmission of IPv6 Packets
%%
\chapter{Transmission of IPv6 Packets over 802.15.4}\label{ch:6lowpan}
As described in Chapter \ref{ch:ieee802.15.4}, the IEEE 802.15.4-2003 standard specifies the physical layer and media access control for low-rate wireless personal area network providing node-to-node frame delivery between the devices within reachable distance from each other. On top of IEEE 802.15.4 a network layer (layer 3 of the seven-layer OSI model) has to be defined that would provide end-to-end packet delivery including routing through intermediate hosts.

>>>>

\chapter{IEEE 802.15.4}\label{ch:ieee802.15.4}
The IEEE 802.15.4 standard \cite{ieee802.15.4} developed by the 802.15.4 Task Group within the IEEE specifies the physical layer and media access control for low-rate wireless personal area networks providing node-to-node frame delivery between devices within reachable distance from each other. The standard is targeted at low cost, low speed and low power consumption devices. This chapter provides an overview of the IEEE 802.15.4 standard.

\section{Device types}
The IEEE 802.15.4 standard distinguishes between two types of nodes, reduced-function devices (RFDs) and full-function devices (FFDs). FFDs typically have more resources, implement the complete protocol set and may be mains powered. An FFD can talk to RFDs or other FFDs, while an RFD can talk only to an FFD. FFDs aid RFDs by providing functions such as network coordination and packet forwarding. An RFD is intended for applications that are extremely simple, which do not have the need to send large amounts of data and may only associate with a single FFD at a time. Therefore, RFDs implement minimal subset of the IEEE 802.15.4 protocol and can be implemented using minimal resources and memory capacity.

Two or more devices within a personal operating space communicating on the same physical channel constitute a wireless personal area network (WPAN). However, a network has to include at least one FFD.

\section{Network topologies}
An 802.15.4 network may operate in either the star or the peer-to-peer topology. In the star topology devices communicate with a single central personal area network (PAN) coordinator. Only an FDD device can be a PAN coordinator. The PAN coordinator is usually mains powered, while the other devices are most likely battery operated. In the peer-to-peer topology a device can communicate with any other device as long as they are in range of each other. Based on the peer-to-peer topology more complex network formations may be constructed, such as mesh networking topology. However, the standard does not define a network layer, therefore, routing is not directly supported, but such an additional layer can add support for multihop communications. All devices operating on a network of either topology have unique 64 bit extended addresses. This address can be used for direct communication within the PAN. Each independent PAN selects a unique identifier. This PAN identifier allows communication between
devices within a network using short 16 bit addresses and enables transmissions between devices across independent networks.

\section{Frame structure}
The basic unit of data transport is a frame. The standard defines four frame structures: a beacon frame, used by a coordinator to transmit beacons; a data frame, used for all transfers of data;  frame, used for confirming successful frame reception; a MAC command frame, used for handling all MAC control transfers. Additionally, a superframe structure may be defined by the coordinator. In such case two beacons act as superframe limits and provide synchronization to other devices. A superframe consists of sixteen equal-length slots, which can be further divided into an active part and an inactive part, during which the coordinator may enter power saving mode. Any device wishing to communicate during the contention access period (CAP) between two beacons shall compete with other devices using a slotted CSMA-CA mechanism. For applications requiring specific data bandwidth, the PAN coordinator may dedicate portions of the active superframe, which form the contention-free period (CFP).

An important aspect of 802.15.4 is its limitation on the frame size, which is specified by the frame length 7 bit field (0-127 bytes). Taking into account the frame header, which is up to 25 octets (disabled security), this leaves 102 bytes for the payload of the higher layers.

\section{Security}
The standard defines several security services such as maintaining an access control list (ACL) and
using symmetric-key cryptography to protect transmitted frames. Access control is a security service that provides the ability for a device to select the other devices with which it is willing to communicate. If the access control service is enabled, a device maintains a list of devices in its ACL from which it expects to receive frames.                        In order to protect data from being read by parties without permission, data may be encrypted using a key shared by a group of devices or using a key shared between two peers. However, key management is not specified in the standard and must be provided by the higher layers. Security services also specify frame integrity that uses a message integrity code to protect data from being modified by parties without the cryptographic key. This provides assurance that data came from a party with the cryptographic key.

%%
%   Transmission of IPv6 Packets
%%
\chapter{Transmission of IPv6 Packets over 802.15.4}\label{ch:6lowpan}
As described in Chapter \ref{ch:ieee802.15.4}, the IEEE 802.15.4-2003 standard specifies the physical layer and media access control for low-rate wireless personal area network providing node-to-node frame delivery between the devices within reachable distance from each other. On top of IEEE 802.15.4, a network layer (layer 3 of the seven-layer OSI model) has to be defined that provides end-to-end packet delivery including routing through intermediate nodes.
